Monaural Single-Source Speaker Separation \\ {\large E511 - Machine Learning for Signal Processing - Final Project}

### 2018-12-15 paper/main.tex
Title update.

% Define document title and author
	\title{Monaural Single-Source Speaker Separation \\
	{\large E511 - Signal Processing - Final Project}}
	\author{Jonathan Branam, Jamie Israel, Jinju(Hellen) Jiang}
	\date{December, 13 2018}
	\maketitle
%
% ^.
% Write abstract here
\begin{abstract}
In this project we evaluated and implemented state of the art techniques to solve the monaural speech separation (diarization) problem using a variety of approached. We investigated classification, masking, and deep learning approaches to separate a generated set of mixed speech recordings.
\end{abstract}

@@

\section{Speech Separation Algorithms}
\subsection{KMeans}
The technique of speaker diarization relies on a big pipeline with following steps:
\begin{itemize}
    \item Feature Extraction
    \item Speaker segmentation
    \item Speaker Clustering
    \item Evaluation
\end{itemize}
\subsubsection{Feature Extraction}
In this project, we used 3 approaches to extract features:
\begin{itemize}
    \item A chroma vector (Wikipedia) that consists of 12 feature vectors indicating how much energy of each pitch class,In the signal we used the same notation as used i Western music notation, to present every pitch class. Then use mean of each element as one of input for clustering.
    \item MFCCs[Wikipedia] are commonly used as features in speech recognition systems, such as the systems which can automatically recognize numbers spoken into a telephone. In this experiment, mfcc computed 13 MFCCs.The first MFCC does not represent information related to the overall shape of the spectrum. It only contains a constant offset. Therefore, here I discard the first MFCC, then I scale the MFCCs such that each coefficient dimension has zero mean and unit variance.Then use mean of MFCCs as one part of input for clustering.
    \item We divide the audio signal into smaller frames. In smaller time scales audio signals are statistically unchanged. Then for each smaller frames we compute the power spectrum of the signal, and use it as the third part of input for clustering.
\end{itemize}
\subsubsection{Speaker segmentation}\cite{Segmentation} Speaker segmentation is to split an audio stream into smaller segments, so as every segment ideally contains only one speaker. So the key point for speech segment is to find out the change point where speakers changed at that point. For example, in our training data, we have conversation between customer service(speaker 0) and customer(speak 1), so we need to find out each change point speaker 0 to speaker 1 or vice versa. In this case, I skipped this step and used the features extracted from above steps as input for kmeans clustering.
\subsubsection{Kmeans Clustering}
Since we knew that the audio from call center is the conversation between 2 speakers(customer service and customer), we set k=2 for kmeans clustering.\\
The results for kmeans speaker diarization are shown in Figure\ref{Fig:speaker diarization 1} and Figure \ref{Fig:speaker diarization 2}.
\begin{figure}[h!]
    \centering  
     \caption{\label{Fig:speaker diarization 1}2 speakers diarization}  
    \includegraphics[width=0.4\textwidth]{kmeans01.png} 
\end{figure}

\begin{figure}[h!]
    \centering  
     \caption{\label{Fig:speaker diarization 2}2 speakers diarization}  
    \includegraphics[width=0.4\textwidth]{kmeans02.png}
\end{figure}
\subsubsection{Evaluation\cite{Segmentation}}The common metric to measure how good/bad for speaker diarization experiments is the Diarization Error Rate(DER), it contains 3 main parts:
\begin{itemize}
    \item The first one is called Missed Speech Rate(MSR), whichi is percentage of scored time that a hypothesized non-speech segment corresponds to a reference speaker segment.
    \item The second one is False alarm speech rate (FASR), which is percentage of scored time that a hypothesized speaker is labelled as a non-speech in the reference.
    \item The third one is Speaker error rate(SER), which is percentage of scored time that a speaker ID is assigned to the wrong speaker.
    \item The last one is Overlap speaker rate(OSR), which is percentage of scored time that some of the multiple speakers in a segment do not get assigned to any speaker.
\end{itemize}

>>>>

% Define document title and author
	\title{Monaural Single-Source Speaker Separation \\
	{\large E511 - Machine Learning for Signal Processing - Final Project}}
	\author{Jonathan Branam, Jamie Israel, Jinju (Hellen) Jiang}
	\date{December, 14 2018}
	\maketitle
%
% ^.
% Write abstract here
\begin{abstract}
In this project we evaluated and implemented state of the art techniques to solve the monaural speech separation (diarization) problem using a variety of approached. We investigated classification, masking, and deep learning approaches to separate a generated set of mixed speech recordings.
\end{abstract}

@@

\section{Speech Separation Algorithms}
\subsection{KMeans}
The technique of speaker diarization relies on a big pipeline with following steps:
\begin{itemize}
    \item Feature Extraction
    \item Speaker segmentation
    \item Speaker Clustering
    \item Evaluation
\end{itemize}
\subsubsection{Feature Extraction}
In this project, we used 3 approaches to extract features:
\begin{itemize}
    \item A chroma vector (Wikipedia) that consists of 12 feature vectors indicating how much energy of each pitch class,In the signal we used the same notation as used i Western music notation, to present every pitch class. Then use mean of each element as one of input for clustering.
    \item MFCCs[Wikipedia] are commonly used as features in speech recognition systems, such as the systems which can automatically recognize numbers spoken into a telephone. In this experiment, mfcc computed 13 MFCCs.The first MFCC does not represent information related to the overall shape of the spectrum. It only contains a constant offset. Therefore, here I discard the first MFCC, then I scale the MFCCs such that each coefficient dimension has zero mean and unit variance.Then use mean of MFCCs as one part of input for clustering.
    \item We divide the audio signal into smaller frames. In smaller time scales audio signals are statistically unchanged. Then for each smaller frames we compute the power spectrum of the signal, and use it as the third part of input for clustering.
\end{itemize}
\subsubsection{Speaker segmentation}\cite{Segmentation} Speaker segmentation is to split an audio stream into smaller segments, so as every segment ideally contains only one speaker. So the key point for speech segment is to find out the change point where speakers changed at that point. For example, in our training data, we have conversation between customer service(speaker 0) and customer(speak 1), so we need to find out each change point speaker 0 to speaker 1 or vice versa. In this case, I skipped this step and used the features extracted from above steps as input for kmeans clustering.
\subsubsection{Kmeans Clustering}
Since we knew that the audio from call center is the conversation between 2 speakers(customer service and customer), we set k=2 for kmeans clustering.\\
The results for kmeans speaker diarization are shown in Figure\ref{Fig:speaker diarization 1} and Figure \ref{Fig:speaker diarization 2}.
\begin{figure}[h!]
    \centering  
     \caption{\label{Fig:speaker diarization 1}2 speakers diarization output}  
    \includegraphics[width=0.4\textwidth]{kmeans01.png} 
\end{figure}

\begin{figure}[h!]
    \centering  
     \caption{\label{Fig:speaker diarization 2}2 speakers diarization on spectrogram}  
    \includegraphics[width=0.4\textwidth]{kmeans02.png}
\end{figure}
\subsubsection{Evaluation\cite{Segmentation}}The common metric to measure how good/bad for speaker diarization experiments is the Diarization Error Rate(DER), it contains 3 main parts:
\begin{itemize}
    \item The first one is called Missed Speech Rate(MSR), whichi is percentage of scored time that a hypothesized non-speech segment corresponds to a reference speaker segment.
    \item The second one is False alarm speech rate (FASR), which is percentage of scored time that a hypothesized speaker is labelled as a non-speech in the reference.
    \item The third one is Speaker error rate(SER), which is percentage of scored time that a speaker ID is assigned to the wrong speaker.
    \item The last one is Overlap speaker rate(OSR), which is percentage of scored time that some of the multiple speakers in a segment do not get assigned to any speaker.
\end{itemize}
